---
output: github_document
---

There are two main components in agua: 

* New parsnip engine `'h2o'` for many models, see the [vignette](https://agua.tidymodels.org/articles/agua.html) for a complete list. 

* Infrastructure for the tune package. 

When fitting a parsnip model, the data are passed to the h2o server directly. For tuning, the data are passed once and instructions are given to `h2o.grid()` to process them. 

This work is based on @stevenpawley's [h2oparsnip](https://github.com/stevenpawley/h2oparsnip) package. Additional work was done by Qiushi Yan for his 2022 summer internship at RStudio. 

## Installation

The CRAN version of the package can be installed via

```r
install.packages("agua")
```

You can also install the development version of agua using:

``` r
require(pak)
pak::pak("tidymodels/agua")
```

## Examples

The following code demonstrates how to create a single model on the h2o server and how to make predictions. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```


```{r}
library(tidymodels)
library(agua)

# Start the h2o server before running models
h2o_start()

# Demonstrate fitting parsnip models: 
# Specify the type of model and the h2o engine 
spec <-
  rand_forest(mtry = 3, trees = 1000) %>%
  set_engine("h2o") %>%
  set_mode("regression")

# Fit the model on the h2o server
set.seed(1)
mod <- fit(spec, mpg ~ ., data = mtcars)
mod

# Predictions
predict(mod, head(mtcars))

# When done
h2o_end()
```

Before using the `'h2o'` engine, users need to run `agua::h2o_start()` or `h2o::h2o.init()` to start the h2o server, which will be storing data, models, and other values passed from the R session. 

There are several package vignettes including: 

- [Introduction to agua](https://agua.tidymodels.org/articles/agua.html)

- [Model tuning](https://agua.tidymodels.org/articles/tune.html)

- [Automatic machine learning](https://agua.tidymodels.org/articles/auto_ml.html)
