---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# agua

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/tidymodels/agua/branch/main/graph/badge.svg)](https://app.codecov.io/gh/tidymodels/agua?branch=main)
[![R-CMD-check](https://github.com/tidymodels/agua/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tidymodels/agua/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

agua enables users to fit, optimize, and evaluate models via [h2o](https://h2o.ai) using a tidymodels interface. 

Most users will not have to use aqua directly; the features can be accessed via a parsnip engine value of `'h2o'`. 

There are two main components in agua: 

* parsnip engine definitions for the following models: 
* Infrastructure for the tune and finetune packages. 

When fitting a parsnip model, the data are passed to the h2o server directly. For tuning, the data are passed once and instructions are given to `h2o.grid()` to process them. 

This work is based on @stevenpawley's [h2oparsnip](https://github.com/stevenpawley/h2oparsnip) package and additional work by Qiushi Yan on his summer internship at RStudio. 

## Installation

You can install the development version of agua like so:

``` r
require(pak)
pak::pak(c("topepo/agua"))
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(tidymodels)
library(agua)

tidymodels_prefer()

# Start the h2o server before running models
h2o::h2o.init()

```

## Code of Conduct
  
Please note that the agua project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
