---
title: "Automatic machine learning"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Automatic machine learning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Using H2O AutoML

Automatic machine learning (AutoML) is the process of automatically searching, screening and evaluating many models for a specific dataset. AutoML could be particularly insightful as an exploratory approach to identify model families and parameterization that is most likely to succeed. You can use H2O's [AutoML](https://docs.h2o.ai/h2o/latest-stable/h2o-docs/automl.html) algorithm via the `'h2o'` engine in `auto_ml()`. agua provides several helper functions to quickly wrangle and visualize AutoML's results.

Let's run an AutoML search on the concrete data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5.75,
  out.width = "95%"
)
options(digits = 3)
```

```{r auto-fit, message = FALSE}
library(tidymodels)
library(agua)
library(ggplot2)
theme_set(theme_bw())
h2o_start()

data(concrete)
set.seed(4595)
concrete_split <- initial_split(concrete, strata = compressive_strength)
concrete_train <- training(concrete_split)
concrete_test <- testing(concrete_split)

# run for a maximum of 120 seconds
auto_spec <-
  auto_ml() %>%
  set_engine("h2o", max_runtime_secs = 120, seed = 1) %>%
  set_mode("regression")

normalized_rec <-
  recipe(compressive_strength ~ ., data = concrete_train) %>%
  step_normalize(all_predictors())

auto_wflow <-
  workflow() %>%
  add_model(auto_spec) %>%
  add_recipe(normalized_rec)

auto_fit <- fit(auto_wflow, data = concrete_train)

extract_fit_parsnip(auto_fit)
```

In 120 seconds, AutoML fitted `r nrow(as.data.frame(auto_fit$fit$fit$fit@leaderboard))` models. The parsnip fit object `extract_fit_parsnip(auto_fit)` shows the number of candidate models, the best performing algorithm and its corresponding model id, and a preview
of the leaderboard with cross validation performances.  The `model_id` column in the leaderboard is a unique model identifier for the h2o server. This can be useful when you need to predict on or extract a specific model, e.g. with `predict(auto_fit, id = id)` and `extract_fit_engine(auto_fit, id = id)`. By default, they will operate on the best performing leader model.

```{r auto-predict}
# predict with the best model
predict(auto_fit, new_data = concrete_test)
```

Typically, we use AutoML to get a quick sense of the range of our success metric, and algorithms that are likely to succeed. agua provides tools to summarize these results.

- `rank_results()` returns the leaderboard in a tidy format with rankings within each metric. A low rank means good performance in a metric. Here, the top 5 models with the smallest MAE includes are four stacked ensembles and one GBM model. 

```{r auto-rank}
rank_results(auto_fit) %>%
  filter(.metric == "mae") %>%
  arrange(rank)
```


- `collect_metrics()` returns average statistics of performance metrics (summarized) per model, or raw value for each resample (unsummarized). `cv_id` identifies the resample h2o internally used for optimization.

```{r auto-metrics}
collect_metrics(auto_fit, summarize = FALSE)
```

- `tidy()` returns a tibble with performance and individual model objects. This is helpful if you want to perform operations (e.g., predict) across all candidates.

```{r auto-tidy}
tidy(auto_fit) %>%
  mutate(
    .predictions = map(.model, predict, new_data = head(concrete_test))
  )
```

- `member_weights()` computes member importance for all stacked ensemble models. Aside from base models such as GLM, GBM and neural networks, h2o tries to fit two kinds of stacked ensembles: one combines the all base models (`"all"`) and the other includes only the best model of each kind (`"bestofFamily"`), specific to a time point. Regardless of how ensembles are formed, we can calculate the variable importance in the ensemble as the importance score of every member model, i.e., the relative contribution of base models in the meta-learner. This is typically the coefficient magnitude in a second-level GLM. This way, in addition to inspecting model performances by themselves, we can find promising candidates if stacking is needed. Here, we show the scaled contribution of different algorithms in stacked ensembles.

```{r auto-stacked}
auto_fit %>%
  extract_fit_parsnip() %>%
  member_weights() %>%
  unnest(importance) %>%
  filter(type == "scaled_importance") %>%
  ggplot() +
  geom_boxplot(aes(value, algorithm)) +
  scale_x_sqrt() +
  labs(y = NULL, x = "scaled importance", title = "Member importance in stacked ensembles")
```


You can also `autoplot()` an AutoML object, which essentially wraps functions above to plot performance assessment and ranking. The lower the average ranking, the more likely the model type suits the data.

```{r autoplot}
autoplot(auto_fit, type = "rank", metric = c("mae", "rmse")) +
  theme(legend.position = "none")
```

After initial assessment, we might be interested to allow more time for AutoML to search for more candidates. Recall that we have set engine argument `max_runtime_secs` to 120s before, we can increase it or adjust `max_models` to control the total runtime. H2O also provides an option to build upon an existing AutoML leaderboard and add more candidates, this can be done via `refit()`. The model to be re-fitted needs to have engine argument `save_data = TRUE`. If you also want to add stacked ensembles set `keep_cross_validation_predictions = TRUE` as well. 

```{r auto-refit, eval = FALSE}
# not run 
auto_spec_refit <-
  auto_ml() %>%
  set_engine("h2o", 
             max_runtime_secs = 300, 
             save_data = TRUE,
             keep_cross_validation_predictions = TRUE) %>%
  set_mode("regression")

auto_wflow_refit <-
  workflow() %>%
  add_model(auto_spec_refit) %>%
  add_recipe(normalized_rec)

first_auto <- fit(auto_wflow_refit, data = concrete_train)
# fit another 60 seconds 
second_auto <- refit(first_auto, max_runtime_secs = 60)
```



## Important engine arguments 

There are several relevant engine arguments for H2O AutoML, some of the most commonly used are: 

- `max_runtime_secs` and `max_models`: Adjust runtime.

- `include_algos` and `exclude_algos`: A character vector naming the algorithms to include or exclude. 

- `validation`: An integer between 0 and 1 specifying the proportion of training data reserved as validation set. This is used by h2o for performance assessment and potential early stopping.


See the details section in `h2o::h2o.automl()` for more information.



## Limiations 

One current limitation of H2O AutoML models is that they can't be used in resampling. This means you can't use them with `fit_resamples()`, `tune_grid()`, `tune_bayes()`, etc.
